---
title: "Exploring Census and Election Data with Microsoft R Open"
author: "Fang Zhou"
date: "Sunday, April 24, 2016"
output: 
    rmarkdown::html_vignette:
        toc: true

vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE}

knitr::opts_chunk$set(fig.width = 6,
                      fig.height = 4,
                      fig.align='center',
                      dev = "png")

```

# Introduction

`eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) is an R package that makes it easy to look at the data from the 2011 Australian Census, and the 2013 Federal Election.

This vignette documents how to explore the 2011 Census and 2013 Federal Election data using Microsoft R Open.

```{r load-read, message=FALSE, error = FALSE}
library(dplyr)
library(ggplot2)
load("data/abs2011.rda")
load("data/aec2013.rda")
```

## 2011 Census Data

Let's look at the data information of 2011 Census.
```{r}
#view data information
rxGetInfo(abs2011,getVarInfo=TRUE)

#transform column type
abs2011<-rxFactors(abs2011,
                   factorInfo=list(
                     Electorate=list(sortLevels=TRUE),
                     State=list(sortLevels=TRUE) 
                                   ),
                   overwrite=TRUE)

rxGetInfo(abs2011,getVarInfo=TRUE)$varInfo[c("Electorate","State")]
```
Let's look at some plots on the census data and the aggregated census data.

## Unemployment
```{r}
#Unemployment
rxHistogram(~Unemployed,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(0,12))
rxHistogram(~Unemployed|State,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(0,12))

tmp<-rxCube(Unemployed~State,abs2011,means=TRUE)
result<-rxResultsDF(tmp)
result
ggplot(result,aes(x=State,y=Unemployed,fill=State))+
  geom_bar(stat="identity")+
  labs(x = "State",
     y = "% Unemployment") + 
  theme_minimal() + 
  theme(legend.position = "none") 
```

## Income
```{r}
#Income
rxHistogram(~MedianIncome,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(250,1100))
rxHistogram(~MedianIncome|State,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(250,1100))

tmp<-rxCube(MedianIncome~State,abs2011,means=TRUE)
result<-rxResultsDF(tmp)
result
ggplot(result,aes(x=State,y=MedianIncome,fill=State))+
  geom_bar(stat="identity")+
  labs(x = "State",
       y = "MedianIncome") + 
  theme_minimal() + 
  theme(legend.position = "none") 
```

## Education
```{r}
#Education
rxHistogram(~Bachelor,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(0,30))
rxHistogram(~Bachelor|State,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(0,30))

tmp<-rxCube(Bachelor~State,abs2011,means=TRUE)
result<-rxResultsDF(tmp)
result
ggplot(result,aes(x=State,y=Bachelor,fill=State))+
  geom_bar(stat="identity")+
  labs(x = "State",
       y = "% electorate with a Bachelor degree") + 
  theme_minimal() + 
  theme(legend.position = "none") 

rxLinePlot(MedianIncome~Bachelor,abs2011,type="p")

tmp<-rxCube(Postgraduate~State,abs2011,means=TRUE)
result<-rxResultsDF(tmp)
result
ggplot(result,aes(x=State,y=Postgraduate,fill=State))+
  geom_bar(stat="identity")+
  labs(x = "State",
       y = "% electorate with a Postgraduate degree") + 
  theme_minimal() + 
  theme(legend.position = "none") 

rxLinePlot(MedianIncome~Postgraduate,abs2011,type="p")

rxLinePlot(Postgraduate+Bachelor~MedianIncome,abs2011,type="p")
```

## Internet
```{r}
#Internet
rxHistogram(~Internet,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(85,100))
rxHistogram(~Internet|State,abs2011,numBreaks=50,histType="Percent",xAxisMinMax=c(85,100))

```

## 2013 Federal Election Data
We have two election datasets. One is called aec2013, and each row represents a particular canditate at a particular polling place. We can and can inspect this data like this:
```{r}
#view data information of 2013 election data
rxGetInfo(aec2013,getVarInfo=TRUE)

#transform column type
aec2013<-rxFactors(aec2013,
                   factorInfo=list(
                     Electorate=list(sortLevels=TRUE),
                     PollingPlace=list(sortLevels=TRUE),
                     Elected=list(sortLevels=TRUE),
                     HistoricElected=list(sortLevels=TRUE),
                     PartyAb=list(sortLevels=TRUE),
                     PartyNm=list(sortLevels=TRUE),
                     State=list(sortLevels=TRUE),
                     DivisionNm.y=list(sortLevels=TRUE)
                                   ),
                   overwrite=TRUE)

rxGetInfo(aec2013,getVarInfo=TRUE)$varInfo[c("PollingPlace","Elected","PartyNm")]

```

We can see that the data contains 82398 rows and 27 columns.

The other dataset is aec2013_electorate and is the results of the election aggregated for each electorate. Each row represents a particular candidate at a particular electorate.

We obtained these data from http://results.aec.gov.au/ . This package incorporates data that is © Commonwealth of Australia (Australian Electoral Commission) 2016.

We can summarise the data to discover some basic details about the election. Here we can find the total number of ordinary votes (not including postal votes, etc.) for each party:

```{r}
tmp<-rxCube(OrdinaryVotes~PartyNm,aec2013,means=FALSE)
total_votes_for_parties<-rxResultsDF(tmp)
library(ggplot2)
library(scales)
ggplot(total_votes_for_parties, aes(reorder(PartyNm, OrdinaryVotes), OrdinaryVotes )) +
  geom_point() + 
  coord_flip() + 
  scale_y_continuous(labels = comma) +
  theme_bw() +
  ylab("Total ordinary votes") +
  xlab("Party") +
  theme(text = element_text(size=10))

```


```{r}
tmp<-rxCrossTabs(OrdinaryVotes~Electorate:PartyNm,aec2013)
result <- rxResultsDF(tmp,output="sums")
result[1:6,1:3]
colnames(result)[-1]<-substring(colnames(result[,-1]),9)
cellnote1<-matrix(as.character(data.matrix(round(result[,-1]))),
                  byrow=FALSE,nrow=dim(result)[1])
library(gplots)
par(oma=c(0.1,0.1,0.1,0.1))
heatmap.2(data.matrix(result[,-1]), 
          labRow=result[,1], col=cm.colors(255),
          trace='none',dendrogram ="none",
          #cellnote=cellnote1,notecex=0.4,notecol="black",
          na.color=par("bg"))

```















